The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-29 paper/paper.tex
A few refinements

\section{Description of iterated parametricity in indexed form\label{sec:reminder}}

@@

\section{Outline of the recursive structure of iterated parametricity construction}

@@

\section{The construction}

@@

\begin{equation*}
  \begin{array}{llcl}
    \depstype[n][\lbrack 0,p-1 \rbrack]    &        & \defeq                               &
    \left(\begin{array}{lll}
              \framep[n][\lbrack 0,p-1 \rbrack]   & : & \frametype[n][\lbrack 0,p-1 \rbrack]    \\
              \painting[n][\lbrack 0,p-1 \rbrack] & : & \paintingtype[n][\lbrack 0,p-1 \rbrack]
              (\framep[n][\lbrack 0,p-1 \rbrack])                                               \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   & : & \restrftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \framep[n][\lbrack 0,p-1 \rbrack]   \\
                  \painting[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                         \\
            \end{array}\right)  \\
    \framep[n+1][\lbrack 0,p \rbrack]      &
    \deps[n][\lbrack 0,p-1 \rbrack]        & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\
    \framep[n+1][\lbrack 0,0 \rbrack]      &
    \unitpoint                             & \defeq & (\unitpoint,\unittype)                 \\
    \framep[n+1][\lbrack 0,p+1 \rbrack]    &
    \deps[n][\lbrack 0,p \rbrack]          & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack]
              (\deps[n][\lbrack 0,p-1 \rbrack]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][\lbrack 0,p-1 \rbrack]).
              \, \layer[n][p]
              (\deps[n][\lbrack 0,p \rbrack])
              (d)                                \\
            \end{array}\right)                                                 \\
    \restrftype[n][\lbrack 0,p-1 \rbrack]  &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
            \end{array}\right) & :      & \U                                                \\
    \restrftype[n]                         &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)                 & \defeq &
    \unittype                                                                                \\
    \restrftype[n][\lbrack 0,p \rbrack]    &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p \rbrack]   \\
              \painting[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)   & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][\lbrack 0,p-1 \rbrack]
      :\restrftype[n][\lbrack 0,p-1 \rbrack]
      \left(\begin{array}{l}
                \framep[n+1][\lbrack 0,p-1 \rbrack]   \\
                \painting[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).         \\
      \Pi q \leq n-p.\,\Pi \epsilon.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][\lbrack 0,p-1 \rbrack]   \\
                \painting[n][\lbrack 0,p-1 \rbrack] \\
                \restrf[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}                       \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \U                                     \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq & \framep[n+1][p]
    (\deps[n][\lbrack 0,p-1 \rbrack])
    \rightarrow \U                                                                           \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \layertype[n][p]
    (\deps[n][\lbrack 0,p \rbrack])
    \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq &
    d \mapsto \Pi\epsilon.\,\painting[n][p](\restrf[n][p][0][\epsilon](d))                   \\
  \end{array}
\end{equation*}

@@

We can now formally define $\painting[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:
\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][p]        &
    \fulldeps[n]             & :      & \paintingtype[n+1][p]   \\
    \painting[n+1][n+1]      &
    \fulldeps[n]             & \defeq & E                       \\
    \painting[n+1][p \leq n] &
    \fulldeps[n]             & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\deps[n][\lbrack 0,p \rbrack]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                   \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

>>>>

\section{High-level description of iterated parametricity in indexed form\label{sec:reminder}}

@@

\section{Analysis of the recursive structure of iterated parametricity in indexed form}

@@

\section{Detailed construction}

@@

\begin{equation*}
  \begin{array}{llcl}
    \depstype[n][\lbrack 0,p-1 \rbrack]    &        & \defeq                               &
    \left(\begin{array}{lll}
              \framep[n][\lbrack 0,p-1 \rbrack]   & : & \frametype[n][\lbrack 0,p-1 \rbrack]    \\
              \painting[n][\lbrack 0,p-1 \rbrack] & : & \paintingtype[n][\lbrack 0,p-1 \rbrack]
              (\framep[n][\lbrack 0,p-1 \rbrack])                                               \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   & : & \restrftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \framep[n][\lbrack 0,p-1 \rbrack]   \\
                  \painting[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                         \\
            \end{array}\right)  \\
    \framep[n+1][\lbrack 0,p \rbrack]      &
    \deps[n][\lbrack 0,p-1 \rbrack]        & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\
    \framep[n+1][\lbrack 0,0 \rbrack]      &
    \unitpoint                             & \defeq & (\unitpoint,\unittype)                 \\
    \framep[n+1][\lbrack 0,p+1 \rbrack]    &
    \deps[n][\lbrack 0,p \rbrack]          & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack]
              (\deps[n][\lbrack 0,p-1 \rbrack]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][\lbrack 0,p-1 \rbrack]).
              \, \layer[n][p]
              (\deps[n][\lbrack 0,p \rbrack])
              (d)                                \\
            \end{array}\right)                                                 \\
    \restrftype[n][\lbrack 0,p-1 \rbrack]  &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
            \end{array}\right) & :      & \U                                                \\
    \restrftype[n]                         &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)                 & \defeq &
    \unittype                                                                                \\
    \restrftype[n][\lbrack 0,p \rbrack]    &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p \rbrack]   \\
              \painting[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)   & \defeq &
    \left\{\begin{array}{l}
      \Sigma \restrf[n][\lbrack 0,p-1 \rbrack]
      :\restrftype[n][\lbrack 0,p-1 \rbrack]
      \left(\begin{array}{l}
                \framep[n+1][\lbrack 0,p-1 \rbrack]   \\
                \painting[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).         \\
      \Pi q \leq n-p.\,\Pi \epsilon.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][\lbrack 0,p-1 \rbrack]   \\
                \painting[n][\lbrack 0,p-1 \rbrack] \\
                \restrf[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}\right.                       \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \U                                     \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq & \framep[n+1][p]
    (\deps[n][\lbrack 0,p-1 \rbrack])
    \rightarrow \U                                                                           \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \layertype[n][p]
    (\deps[n][\lbrack 0,p \rbrack])
    \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq &
    d \mapsto \Pi\epsilon.\,\painting[n][p](\restrf[n][p][0][\epsilon](d))                   \\
  \end{array}
\end{equation*}

@@

We can now formally define $\painting[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:
\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][p]        &
    \fulldeps[n]             & :      & \paintingtype[n+1][p]   \\
    \painting[n+1][n+1]      &
    \fulldeps[n]             & \defeq & E                       \\
    \painting[n+1][p \leq n] &
    \fulldeps[n]             & \defeq & \lambda d.\,
              \Sigma l:\layer[n][p](\deps[n][\lbrack 0,p \rbrack]). 
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                   \\
  \end{array}
\end{equation*}
